Associative Data Structures for Website Filtering

### 2014-04-17 paper/paper.tex
Tried to make things more formal.

    \begin{abstract}
        In this paper, we will research associative data structures for searching. This is done in the context of
        a website filter. We look at three different data structures: the hash map, the binary search tree
        and the skip list. We insert keys of three different types, namely IP adress, domain name and
        full path URL. We judge the competency of our data strucutre based on insertion times, query times
        and memory usage, inserting data sets of sizes $2^x$ for $8 \le x \le 20$. We describe the details
        of our research and the implementation using C++. The results are that hash maps are much faster
        than the other structures, skip lists use the least memory and binary search trees are
        only useful with very few insertions of simple keys and more insertions than queries.
    \end{abstract}

    \section{Introduction}
    A website filter is a program that categorises websites in order to restrict access to designated
    sites a user might visit. Commonly, the filter includes a blacklist of sites a user may not visit
    unless they have certain credentials, or alternatively, includes a whitelist so it can blacklist
    everything else. As the world-wide web is growing, these lists have grown larger to a point where
    the speed of the underlying data structure has become a key issue. The time of both inserting and
    looking up websites must be kept minimal. This all must be done without too much of a trade-off
    between speed and memory usage.

    In technical terms, we would want the best data structure for associating many URLs with their
    respective permissions, be it whitelisted or blacklisted for certain users. In researching the best
    associative data structure we have picked implementations of the binary search tree, the hash map
    and the skip list. These are the key options in such a problem. We are interested in finding which
    one would perform best under a various number of insertions. The main research question of this paper
    is as follows:

    \begin{quotation}How does the number of entries influence the performance of various associative data structures?\end{quotation}

@@

    \begin{quotation}How does the complexity of the key influence the preformance of various associative data structures?\end{quotation}

    In the following paper, we will attempt to answer these questions. The next section will provide details
    about our research plan and the scopes of our research. The section thereafter is about the implementation
    of our setup in C++. Next, we will state our results, after which we will discuss set results and provide
    several conclusions.    
    %We did research about what data structure could best be used to represent a blacklist in a 
    %web-filter application.  We compared the insertion speed, query speed, and memory usage of three
    %data structures: the binary search tree, the hash table, and the skip list.  We did this because
    %we really liked these three structures. 
    \section{Implementation}
    \subsubsection*{Datastructures}

@@

    \subsubsection*{Data}
    For our data, we took the system's dictionary file and the list of top level domains from Wikipedia
    and combined these to generate a list of items in the form \texttt{www.\$word.\$tld}, where
    \texttt{\$word} is one of words from the the dictionary and \texttt{\$tld} is a top level domain.  Only
    words consisting of the lowercase letters \texttt{a} through \texttt{z} were used.  Constraining the
    words to consist of 8 to 12 characters and limiting the top-level domains to exactly two letters, 
    all possible combinations were generated.  For the full paths, a number of domains were selected
    and each was appended a random alphanumeric string of eighty characters.  The exact implementation
    can be viewed in the sourcecode\footnote{ref to apendix or code dump}.

	\subsubsection*{Measurements}
    For our time measurements, we used libstdc++'s implementation of the standard
    \texttt{std::chrono::high\_resolution\_clock} class.

>>>>

    \begin{abstract}
        We have researched the performance of the hash table, binary search tree, and skip list
        associative datastructures in the context of a web filter.  Our measurements were made with
        structures of various sizes (ranging from $2^8$ up to $2^{20}$) and with different possible
        key types (from 4 to 120 bytes).  In this paper, we describe our implementation and
        empirically show that given a sufficiently high number of entries, a hash table has the
        highest insertion and query speeds, while the skip list requires the least memory per
        element.
        % And when I say "hash table", I mean libstdc++'s unordered_map, and when I say bst I mean
        % libstdc++'s map, and if you took different implementations you'd probably get different
        % results...
    \end{abstract}

    \section{Introduction}
    A website filter is a program that categorises websites in order to restrict access to designated
    sites a user might visit. Commonly, the filter operates by keeping a blacklist of sites a user may not visit
    unless they have certain credentials, or alternatively, keeping a whitelist and prohibiting any
    sites not on it.  As the world-wide web is growing, these lists have grown larger to a point where
    the speed of the underlying data structure has become a key issue.  The time of both inserting and
    looking up websites must be kept minimal. Moreover, the increased number of entries make memory
    usage an equally significant concern.

    In technical terms, we are interested in the best data structure for associating many URLs with
    their respective permissions, be it whitelisted or blacklisted for certain users.  We have
    chosen to compare the performance of the binary search tree, the hash map, and the skip list, as
    these are commonly employed in practice.  It is of particular interest how each data structure
    will scale as the number of elements rises.  We have thus chosen for our main research question
    the following:
    % Guys, did you notice we did absolutely nothing about predicting this?

    \begin{quotation}
        How does the number of entries influence the performance of various associative data
        structures?
    \end{quotation}

@@

    \begin{quotation}
        How does the complexity of the key influence the preformance of various associative data structures?
    \end{quotation}

    In this paper, we attempt to answer these questions. Section~\ref{sec:plan} provides details
    about our research plan and the scopes of our research. Section~\ref{sec:implementation} is
    about the implementation of our setup in C++.  The rest of the paper states our results and the
    conclusions that can be drawn from them.

    \section{Plan}
    \label{sec:plan}

    \section{Implementation}
    \label{sec:implementation}

    We used C++ implementations of the data structures involved, and also the language for our
    measuring code.  We chose C++ as it is an industry standard\footnote{ref}, provides the
    necessary tools, and does not carry the risk of garbage collection pauses or effects due to
    ``warming-up time''.

    \subsection*{Data Structures}

@@

    \subsection*{Data}
    For our data, we took the system's dictionary file and the list of top level domains from Wikipedia
    and combined these to generate a list of items in the form \texttt{www.\$word.\$tld}, where
    \texttt{\$word} is one of words from the the dictionary and \texttt{\$tld} is a top level domain.  Only
    words consisting of the lowercase letters \texttt{a} through \texttt{z} were used.  Constraining the
    words to consist of 8 to 12 characters and limiting the top-level domains to exactly two letters, 
    all possible combinations were generated.  For the full paths, a number of domains were selected
    and each was appended a random alphanumeric string of eighty characters.  The exact implementation
    can be viewed in the sourcecode\footnote{ref to apendix or code dump}.

    \subsection*{Measurements}
    For our time measurements, we used libstdc++'s implementation of the standard
    \texttt{std::chrono::high\_resolution\_clock} class.
